Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-15 main.Rmd
finer, coarser and bottom granularity defined informally

Discrete abstraction of time such as hours, days, weeks and months can be thought of as linear "time granularities" [@aigner2011visualization]. Time granularities are **linear** if they respect the linear progression of time.

@@

A discrete time model often uses a fixed smallest granularity also known as _bottom granularity_ (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn). Chronons may be grouped into granules and a non-overlapping mapping of granules to subsets of the time domain is a linear granularity. A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.

<!-- CONTENT CUT -->
<!-- For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$. For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$. The _altering-tick operation_ generates new full-integer labeled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$. For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf. -->

@@

The altering-tick operation:
: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$. The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ generates a new full-integer labeled granularity by periodically expanding or shrinking granules of $G_1$ in terms of granules of $G_2$. The altering-tick operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each group has $k$ additional granules of $G_2$.

>>>>

Discrete abstraction of time such as weeks, months or holidays can be thought of as "time granularities". Time granularities are **linear** if they respect the linear progression of time.

@@

A discrete time model often uses a fixed smallest linear granularity also known as _bottom granularity_ (@Bettini1998-ed) <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn). Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity. A system of multiple granularities in lattice structures is referred to as a _calendar_ (see @Dyreson2000-mr).

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.

<!-- CONTENT CUT -->
<!-- For a coarser granularity $G_2$ and a finer one $G_1$, a _grouping operation_ partitions some granules of $G_1$  to form a granule of $G_2$. For example, $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$. The _altering-tick operation_ generates new full-integer labeled granularity by periodically expanding or shrinking granules of $G_2$ in terms of granules of $G_1$. For example, <!--due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf.

@@

The altering-tick operation:
: Let $G_1$, $G_2$ be full-integer labeled granularities, and $l$, $k$, $m$ integers, where $G_2$ partitions $G_1$, and $1 \leq l \leq m$. The altering-tick operation $\text{Alter}^{m}_{l,k} (G_2, G_1)$ operation modifies the granules of $G_1$ so that the $l^{\text{th}}$ granule of each period has $k$ additional granules of $G_2$.
